Climate indicators

### 2023-02-23 DRAFT_climate_indicators.Rmd
break points scaling

# Function to add random noise
myRandom <- function(x) x*rnorm(1,1,.1)

@@

dim(y1)
dim(y2)

@@

```{r, eval=F, include=F}
# I would normally do the above code chink using tmap, but I cant get it to work when faceting across dimensions:
tm_shape(y)+
  tm_raster()+
  tm_facets("time")
```

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, more or less divide them by two. I then merge with the reference value and sd, creating one common data cube.

```{r}
y_var <- y %>% mutate(Attribute_A = Attribute_A/2) %>% 
  split(3) %>% c(y_ref)
y_var <- y_var %>%
  setNames(paste0("v_",substr(names(y_var), 1,4)))%>%
  rename("reference" = v_refe,
         "sd" = v_sd)
```

```{r}
# select the columns to normalise
cols <- names(y_var)[!names(y_var) %in% c("reference", "sd") ]
cols_new <- cols
names(cols_new) <- gsub("v_", "i_", cols)
# Mutate 
y_var_norm <- y_var %>%
  mutate(across(all_of(cols), ~.x / reference)) %>%
  rename(all_of(cols_new))

#y_v <- c(y1v, select(y_ref, reference_value))
#y_v_merged <- merge(y_v)
#dim(y_v_merged)
#names(y_v_merged)
#y_v_merged <- setNames(y_v_merged, "value")
#st_get_dimension_values(y_v_merged, "attributes")
```

>>>>

# Function to add random noise
myRandom <- function(x) x*rnorm(1,1,.05)

@@

@@

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, taking them down 2%. I then merge with the reference value and sd, creating one common data cube.

```{r}
y_var <- y %>% mutate(Attribute_A = Attribute_A * 0.98) %>% 
  split(3) %>% c(y_ref)
y_var <- y_var %>%
  setNames(paste0("v_",substr(names(y_var), 1,4)))%>%
  rename("reference" = v_refe,
         "sd" = v_sd)
```

```{r}
# select the columns to normalise
cols <- names(y_var)[!names(y_var) %in% c("reference", "sd") ]
cols_new <- cols
names(cols_new) <- gsub("v_", "i_", cols)

# Mutate
# can I find a way to keep the v_ columns?
y_var_norm <- y_var %>%
  mutate(lower = reference - 5*sd ) %>%
  mutate(threshold = reference -2*sd ) %>%
  mutate(across(all_of(cols), ~ if_else(.x < threshold, 
                                        (.x - lower) / (threshold - lower),
                                        (.x - threshold) / (reference - threshold),
                                        ))) %>%
  mutate(across(all_of(cols), ~ if_else(.x > 1, 1, .x))) %>%
  mutate(across(all_of(cols), ~ if_else(.x < 0, 0, .x))) %>%
  rename(all_of(cols_new))
```
